### 2016-07-07 responseToReviewers.Rmd

---
output:
  word_document:
    fig_caption: true
  pdf_document:
    fig_caption: true
    latex_engine: xelatex
    keep_tex: yes
header-includes:
   - \usepackage{booktabs}
   - \usepackage[final]{changes}
   - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
   - \linespread{1.2}
   - \usepackage[compact]{titlesec}
   - \usepackage{enumitem}
   - \usepackage{tikz}
   - \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
   - \setlist{nolistsep}
   - \titlespacing{\section}{2pt}{*0}{*0}
   - \titlespacing{\subsection}{2pt}{*0}{*0}
   - \titlespacing{\subsubsection}{2pt}{*0}{*0}
   - \setlength{\parskip}{3pt}
   - \setremarkmarkup{(#2)}
bibliography: references.bib
csl: national-science-foundation-grant-proposals.csl
fontsize: 12pt
mainfont: Georgia
---

@@

> _\textcolor{blue}{For this particular application, white matter hyperintensities
are the structure of interest.  We have re-emphasized this point by writing:
}_

> \textcolor{blue}{Crucial to these supervised segmentation approaches are the creation and selection of
``features'' as input (i.e., feature images constructed from the training data)
in conjunction with expertly identified structures of interest
(i.e., WMHs) for model construction.}

6. The methods section is very confusing. The section "Feature images for
WMH segmentation" starts with no explanation of what is Stage 1. The
logical flow would require template registration explained first, and all
feature computation performed later, and finally RF stages can be
explained with the available features.

> _\textcolor{blue}{We added Figure 1 which gives an overview of the basic workflow.
>   It can be seen below in response to item 2 of Reviewer 2.  We reference
>   this figure numerous times in the text as a visual explanatory aid.}_
> _\textcolor{blue}{We also added a paragraph at the beginning of the "Quantitative analysis" subsection
>   which gives an overview of the methods section including an description of Stage 1 and Stage2.}_
>
>   \textcolor{blue}{Figure 1 provides a graphical overview of the proposed workflow.  The major components
include offline generation of symmetric multimodal templates, the creation of
feature images from the training data which are then employed for statistical
prediction using a concatenated random forest framework.  This framework involves
the use of two random forest models where the outcome (i.e., the tissue
probability estimates) of the first RF model application,
which we denote as ``Stage 1'', is used as input (along with the original set of
feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
Once these offline steps
are performed, a new (i.e., unsegmented) subject can then be processed using the proposed pipeline.}
>
> _\textcolor{blue}{We also added a subsection explaining the symmetric multimodal template:}_
>
> \textcolor{blue}{Following} [@Tustison:2014ab] \textcolor{blue}{and} [@Tustison:2015aa], \textcolor{blue}{optimally derived templates
> serve for both brain tissue segmentation (for the derivation of tissue prior probability maps) and generation of asymmetry feature images.
> For this work we use the multi-modal data available from the public MMRR data set}
> [@landman2011]\textcolor{blue}{.  We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in} [@Avants:2010aa] \textcolor{blue}{which
> results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.}
>
> _\textcolor{blue}{These changes complement the edits made in response to other review items.}_

@@

> _\textcolor{blue}{Done.  We removed the mathematical formulas.}_

7. Random forests are explained in a way that can be understood only by
readers who already know how they work; a new reader (especially a
clinician) would not be able to understand the method. For example, data
pushed onto a tree is does not mean much, unless the reader knows what is
an identification tree.

> _\textcolor{blue}{
We revised the explanation of random forests as follows:}_

> \textcolor{blue}{The basic component of
the random forest paradigm is the ``decision tree'' often represented by a flowchart or
graph where internal nodes represent ``tests'', or decisions, and the edges represent the outcome of those tests.  The final, or end, nodes represent the various classifications produced by traversal through the decision tree.  For the proposed application, individual
voxels (and their corresponding feature values) are introduced at the root of a particular
decision tree and traverse the edges and internal nodes ultimately ending up at one of
the classification nodes according to the tests at each internal node.  A single random
forest model will consist of many such trees (often refered to as an ``ensemble'').}
> \textcolor{blue}{Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
(e.g., AdaBoost)} [@schapire1990;@freund1997]
>  \textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization} [@ho1995;@amit1997].
\textcolor{blue}{Finally, Breiman} [@breiman2001]
\textcolor{blue}{improved accuracy by random sampling of training data (i.e., ``bagging'') resulting
in the current random forest technique applied here.}
> \textcolor{blue}{As voxels and their feature values are ``pushed'' through each decision tree in the forest,  votes for each label are accumulated and converted to probability values for all classification possiblilities at
each voxel location.}

@@

Interesting paper which offers those without a current background in
state-of-the-art brain analysis tools. A good introduction to their capabilities.

> _\textcolor{blue}{We thank the reviewer for the positive comments regarding our work.}_

1. However the authors do little to establish clinical utility of the
algorithm.

> _\textcolor{blue}{We appreciate
> the opportunity to clarify the precise nature of our contribution which is methodological
> (vs. clinical).  We have shortened the introduction to focus specifically on previous TBI
clinical research involving WMHs and then point to the need for automated
techniques to minimize the variability and time requirements of manual
segmentation of WMHs.  This would allow for large-scale studies and potentially more accurate
and reproducible quantitation.
}_

### 2016-07-07 responseToReviewers.tex

\begin{quote}
\emph{\textcolor{blue}{We completely reorganized the results and discussion
section to reflect the new evaluation.  In addition, we added a synopsis
of these results to the abstract.}}
\end{quote}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Third, there is no information about the manual lesion tracing
  procedure. Were lesion drawn on FLAIR or were other modalities used as
  well? If FLAIR was used as reference it is not surprising that FLAIR
  emerges as one of the top predictive modalities.
\end{enumerate}

\begin{quote}
\emph{\textcolor{blue}{We added the following to the Methods section:}}
\end{quote}

\begin{quote}
\textcolor{blue}{The first author (J. R. S.) performed the manual WMH tracings for all 26  subjects.  J. R. S. is a radiologist certified by the American Board of Radiology, with a certificate of advanced qualification in vascular and interventional radiology, over 18 years of research experience in TBI, and 6 years of clinical imaging experience.  All multi-modal MR dicom image slices were converted to the nifti file format. All nifti image volumes for each subject were rigidly aligned to the T1 image of that subject using the ANTs software}
{[}1{]}
\textcolor{blue}{.  The normalized MRI volumes were then provided to J. R. S. who traced each lesion using the ITK-SNAP tool}
{[}2{]}
\textcolor{blue}{ which has multi-image overlay capabilities for visualizing all modalities in all three canonical views.}
\end{quote}

@@

\begin{quote}
\textcolor{blue}{Crucial to these supervised segmentation approaches are the creation and selection of
``features'' as input (i.e., feature images constructed from the training data)
in conjunction with expertly identified structures of interest
(i.e., WMHs) for model construction.}
\end{quote}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{5}
\tightlist
\item
  The methods section is very confusing. The section ``Feature images
  for WMH segmentation'' starts with no explanation of what is Stage 1.
  The logical flow would require template registration explained first,
  and all feature computation performed later, and finally RF stages can
  be explained with the available features.
\end{enumerate}

\begin{quote}
\emph{\textcolor{blue}{We added Figure 1 which gives an overview of the basic workflow.
  It can be seen below in response to item 2 of Reviewer 2.  We reference
  this figure numerous times in the text as a visual explanatory aid.}}
\emph{\textcolor{blue}{We also added a paragraph at the beginning of the "Quantitative analysis" subsection
  which gives an overview of the methods section including an description of Stage 1 and Stage2.}}

\textcolor{blue}{Figure 1 provides a graphical overview of the proposed workflow.  The major components
include offline generation of symmetric multimodal templates, the creation of
feature images from the training data which are then employed for statistical
prediction using a concatenated random forest framework.  This framework involves
the use of two random forest models where the outcome (i.e., the tissue
probability estimates) of the first RF model application,
which we denote as ``Stage 1'', is used as input (along with the original set of
feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
Once these offline steps
are performed, a new (i.e., unsegmented) subject can then be processed using the proposed pipeline.}

@@

\emph{\textcolor{blue}{We also added a subsection explaining the symmetric multimodal template:}}

\textcolor{blue}{Following} {[}3{]} \textcolor{blue}{and} {[}4{]},
\textcolor{blue}{optimally derived templates
serve for both brain tissue segmentation (for the derivation of tissue prior probability maps) and generation of asymmetry feature images.
For this work we use the multi-modal data available from the public MMRR data set}
{[}5{]}\textcolor{blue}{.  We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in}
{[}6{]} \textcolor{blue}{which
results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.}

@@

\emph{\textcolor{blue}{These changes complement the edits made in response to other review items.}}
\end{quote}

\begin{quote}
The pipeline also seem to be repetitive, the N4 correction is mentioned
in ln. 26 and then again in ln. 43, which sounds like it was performed
twice. Tissue segmentation is mentioned in ln. 39, then it is stated is
a Bayesian method (ln 45), and then it is referred to as Atropos in the
next page.
\end{quote}

\begin{quote}
\emph{\textcolor{blue}{The two mentions of N4 are due to the fact that they are associated
with two distinct pipelines:  1) preparation of the images for extracting feature values
and 2) brain extration and $n$-tissue segmentation.  We rewrote the section in question
to avoid any confusing redundancy:}}
\end{quote}

\begin{quote}
\emph{\textcolor{blue}{The T1 image is then processed via the ANTs brain
extraction and normal tissue segmentation pipelines}} {[}3{]}
\emph{\textcolor{blue}{.  The result is a mask delineating the brain
parenchyma and probabilistic estimates of the CSF, gray matter, white matter,
deep gray matter, brain stem, and cerebellum}} {[}7{]}
\emph{\textcolor{blue}{.  These provide the expertly annotated labels for the
first six tissue labels given above.  Tissue prior probability maps for segmentation
are from multi-model optimal symmetric shape/intensity templates}}
{[}4{]} \emph{\textcolor{blue}{created from the public MMRR data set}}
{[}5{]} \emph{\textcolor{blue}{(cf Figure 3).}}
\end{quote}

@@

\begin{quote}
\emph{\textcolor{blue}{Done.  We removed the mathematical formulas.}}
\end{quote}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{6}
\tightlist
\item
  Random forests are explained in a way that can be understood only by
  readers who already know how they work; a new reader (especially a
  clinician) would not be able to understand the method. For example,
  data pushed onto a tree is does not mean much, unless the reader knows
  what is an identification tree.
\end{enumerate}

\begin{quote}
\emph{\textcolor{blue}{
We revised the explanation of random forests as follows:}}
\end{quote}

\begin{quote}
\textcolor{blue}{The basic component of
the random forest paradigm is the ``decision tree'' often represented by a flowchart or
graph where internal nodes represent ``tests'', or decisions, and the edges represent the outcome of those tests.  The final, or end, nodes represent the various classifications produced by traversal through the decision tree.  For the proposed application, individual
voxels (and their corresponding feature values) are introduced at the root of a particular
decision tree and traverse the edges and internal nodes ultimately ending up at one of
the classification nodes according to the tests at each internal node.  A single random
forest model will consist of many such trees (often refered to as an ``ensemble'').}
\textcolor{blue}{Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
(e.g., AdaBoost)} {[}8, 9{]}
\textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization}
{[}10, 11{]}. \textcolor{blue}{Finally, Breiman} {[}12{]}
\textcolor{blue}{improved accuracy by random sampling of training data (i.e., ``bagging'') resulting
in the current random forest technique applied here.}
\textcolor{blue}{As voxels and their feature values are ``pushed'' through each decision tree in the forest,  votes for each label are accumulated and converted to probability values for all classification possiblilities at
each voxel location.}
\end{quote}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{7}
\tightlist
\item
  The relationship between WMH and GOS-E is mentioned in pg. 2 but the
  reader is not informed what is GOS-E. A similar issue is met few lines
  after, where the ``outcome'' is mentioned, but there is no information
  what is referred to as outcome.
\end{enumerate}

\begin{quote}
\emph{\textcolor{blue}{The sentence in question reads:}}
\end{quote}

\begin{quote}
\textcolor{blue}{Further, volume of FLAIR lesions within the corpus
callosum, brainstem, and thalamus in patients with severe TBI correlates
with Glasgow Outcome-Extended (GOS-E) scores.} {[}13{]}
\end{quote}

@@

\begin{quote}
\emph{\textcolor{blue}{Although the full phrase corresponding to the acronym
in question is provided prior to the acronym in the text (along with an
explanatory reference), we have added the following clarification:}}
\end{quote}

\begin{quote}
\textcolor{blue}{Further, volume of FLAIR lesions within the corpus
callosum, brainstem, and thalamus in patients with severe TBI correlates
with Glasgow Outcome-Extended (GOS-E) scores} {[}13{]}
\textcolor{blue}{---a numeric groupwise assessment used to
classify ``outcome'' in TBI patients where ``outcome'' refers to the
spectrum of possible prognoses from death to disability to recovery.}
\end{quote}

@@

\begin{quote}
\emph{\textcolor{blue}{This was a grammatical mistake, not a question of
ethics.  It has been fixed: ``The feature images
were derived from MR acquisitions of 26 subjects\ldots''}}
\end{quote}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{10}
\tightlist
\item
  It is unclear to me the meaning of the sentence: ``Since the inverse
  transform is also derived as part of the registration process, we can
  warp the voxel index locations back to the space of the individual
  subject which motivates similar work by others {[}43{]}.''
\end{enumerate}

\begin{quote}
\emph{\textcolor{blue}{The sentences were rewritten as follows:}}
\end{quote}

\begin{quote}
\textcolor{blue}{Similar feature images were used in} {[}14{]}
\textcolor{blue}{although, unlike the proposed framework, this previous work lacks
normalization to the standard
coordinate system provided by the template to dramatically improve spatial specificity
across all subjects.  To generate these images, the T1 image of each subject is
registered to the T1 template component using a B-spline variant}
{[}15{]}
\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm}
{[}16{]}.
\textcolor{blue}{Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.}
\end{quote}

@@

Interesting paper which offers those without a current background in
state-of-the-art brain analysis tools. A good introduction to their
capabilities.

\begin{quote}
\emph{\textcolor{blue}{We thank the reviewer for the positive comments regarding our work.}}
\end{quote}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  However the authors do little to establish clinical utility of the
  algorithm.
\end{enumerate}

\begin{quote}
\emph{\textcolor{blue}{We appreciate
the opportunity to clarify the precise nature of our contribution which is methodological
(vs. clinical).  We have shortened the introduction to focus specifically on previous TBI
clinical research involving WMHs and then point to the need for automated
techniques to minimize the variability and time requirements of manual
segmentation of WMHs.  This would allow for large-scale studies and potentially more accurate
and reproducible quantitation.
}}
\end{quote}

@@

\begin{figure}[htbp]
\centering
\includegraphics{Figures/wmhPipeline.png}
\caption{\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
input MRI for a single subject, using the multi-modal symmetric template, results in
the generation of the feature images.  These feature images are used as input to the
Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
voting maps, the original feature images, and the Stage 2 RF model result in the
final voting maps which includes the WMH probability estimate.  Note that the RF models
are constructed once from a set of training data which are processed using the
same feature-construction pipeline as the single-subject input MRI.}}
\end{figure}

\clearpage

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\hypertarget{ref-Avants:2014aa}{}
1. Avants, B. B., Tustison, N. J., Stauffer, M., Song, G., Wu, B., and
Gee, J. C. ``\textbf{The Insight ToolKit Image Registration Framework}''
\emph{Front Neuroinform} 8, (2014): 44.
doi:\href{https://doi.org/10.3389/fninf.2014.00044}{10.3389/fninf.2014.00044}

\hypertarget{ref-Yushkevich:2006aa}{}
2. Yushkevich, P. A., Piven, J., Hazlett, H. C., Smith, R. G., Ho, S.,
Gee, J. C., and Gerig, G. ``\textbf{User-Guided 3D Active Contour
Segmentation of Anatomical Structures: Significantly Improved Efficiency
and Reliability}'' \emph{Neuroimage} 31, no. 3 (2006): 1116--28.
doi:\href{https://doi.org/10.1016/j.neuroimage.2006.01.015}{10.1016/j.neuroimage.2006.01.015}

\hypertarget{ref-Tustison:2014ab}{}
3. Tustison, N. J., Cook, P. A., Klein, A., Song, G., Das, S. R., Duda,
J. T., Kandel, B. M., Strien, N. van, Stone, J. R., Gee, J. C., and
Avants, B. B. ``\textbf{Large-Scale Evaluation of ANTs and FreeSurfer
Cortical Thickness Measurements}'' \emph{Neuroimage} 99, (2014):
166--79.
doi:\href{https://doi.org/10.1016/j.neuroimage.2014.05.044}{10.1016/j.neuroimage.2014.05.044}

\hypertarget{ref-Tustison:2015aa}{}
4. Tustison, N. J., Shrinidhi, K. L., Wintermark, M., Durst, C. R.,
Kandel, B. M., Gee, J. C., Grossman, M. C., and Avants, B. B.
``\textbf{Optimal Symmetric Multimodal Templates and Concatenated Random
Forests for Supervised Brain Tumor Segmentation (Simplified) with
ANTsR}'' \emph{Neuroinformatics} 13, no. 2 (2015): 209--25.
doi:\href{https://doi.org/10.1007/s12021-014-9245-2}{10.1007/s12021-014-9245-2}

\hypertarget{ref-landman2011}{}
5. Landman, B. A., Huang, A. J., Gifford, A., Vikram, D. S., Lim, I. A.
L., Farrell, J. A. D., Bogovic, J. A., Hua, J., Chen, M., Jarso, S.,
Smith, S. A., Joel, S., Mori, S., Pekar, J. J., Barker, P. B., Prince,
J. L., and Zijl, P. C. M. van. ``\textbf{Multi-Parametric Neuroimaging
Reproducibility: A 3-T Resource Study}'' \emph{Neuroimage} 54, no. 4
(2011): 2854--66.
doi:\href{https://doi.org/10.1016/j.neuroimage.2010.11.047}{10.1016/j.neuroimage.2010.11.047}

\hypertarget{ref-Avants:2010aa}{}
6. Avants, B. B., Yushkevich, P., Pluta, J., Minkoff, D., Korczykowski,
M., Detre, J., and Gee, J. C. ``\textbf{The Optimal Template Effect in
Hippocampus Studies of Diseased Populations}'' \emph{Neuroimage} 49, no.
3 (2010): 2457--66.
doi:\href{https://doi.org/10.1016/j.neuroimage.2009.09.062}{10.1016/j.neuroimage.2009.09.062}

\hypertarget{ref-Avants:2011aa}{}
7. Avants, B. B., Tustison, N. J., Wu, J., Cook, P. A., and Gee, J. C.
``\textbf{An Open Source Multivariate Framework for \(n\)-Tissue
Segmentation with Evaluation on Public Data}'' \emph{Neuroinformatics}
9, no. 4 (2011): 381--400.
doi:\href{https://doi.org/10.1007/s12021-011-9109-y}{10.1007/s12021-011-9109-y}

\hypertarget{ref-schapire1990}{}
8. Schapire, R. ``\textbf{The Strength of Weak Learnability}''
\emph{Machine Learning} 5, (1990): 197--227.

\hypertarget{ref-freund1997}{}
9. Freund, Y. and Schapire, R. ``\textbf{A Decision-Theoretic
Generalization of on-Line Learning and an Application to Boosting}''
\emph{Journal of Computer and System Sciences} 55, (1997): 119--139.

\hypertarget{ref-ho1995}{}
10. Ho, T. K. ``\textbf{Random Decision Forests}'' \emph{Document
analysis and recognition, 1995., proceedings of the third international
conference on} 1, (1995): 278--282 vol.1.
doi:\href{https://doi.org/10.1109/ICDAR.1995.598994}{10.1109/ICDAR.1995.598994}

\hypertarget{ref-amit1997}{}
11. Amit, Y. and Geman, D. ``\textbf{Shape Quantization and Recognition
with Randomized Trees}'' \emph{Neural Computation} 9, (1997):
1545--1588.

\hypertarget{ref-breiman2001}{}
12. Breiman, L. ``\textbf{Random Forests}'' \emph{Machine learning}
(2001): 5--32.

\hypertarget{ref-Moen:2014aa}{}
13. Moen, K. G., Brezova, V., Skandsen, T., Håberg, A. K., Folvik, M.,
and Vik, A. ``\textbf{Traumatic Axonal Injury: The Prognostic Value of
Lesion Load in Corpus Callosum, Brain Stem, and Thalamus in Different
Magnetic Resonance Imaging Sequences}'' \emph{J Neurotrauma} 31, no. 17
(2014): 1486--96.
doi:\href{https://doi.org/10.1089/neu.2013.3258}{10.1089/neu.2013.3258}

\hypertarget{ref-Anbeek:2004aa}{}
14. Anbeek, P., Vincken, K. L., Osch, M. J. P. van, Bisschops, R. H. C.,
and Grond, J. van der. ``\textbf{Probabilistic Segmentation of White
Matter Lesions in MR Imaging}'' \emph{Neuroimage} 21, no. 3 (2004):
1037--44.
doi:\href{https://doi.org/10.1016/j.neuroimage.2003.10.012}{10.1016/j.neuroimage.2003.10.012}

\hypertarget{ref-Tustison:2013ac}{}
15. Tustison, N. J. and Avants, B. B. ``\textbf{Explicit B-Spline
Regularization in Diffeomorphic Image Registration}'' \emph{Front
Neuroinform} 7, (2013): 39.
doi:\href{https://doi.org/10.3389/fninf.2013.00039}{10.3389/fninf.2013.00039}

\hypertarget{ref-Avants:2011ab}{}
16. Avants, B. B., Tustison, N. J., Song, G., Cook, P. A., Klein, A.,
and Gee, J. C. ``\textbf{A Reproducible Evaluation of ANTs Similarity
Metric Performance in Brain Image Registration}'' \emph{Neuroimage} 54,
no. 3 (2011): 2033--44.
doi:\href{https://doi.org/10.1016/j.neuroimage.2010.09.025}{10.1016/j.neuroimage.2010.09.025}

### 2016-07-07 abstract.Rmd
ENH:  Second round.

White matter hyperintensities (WMHs) are foci of abnormal signal
intensity in white matter regions seen with magnetic resonance imaging
(MRI). WMHs are associated with normal aging and have shown prognostic
value in neurological conditions such as traumatic brain injury (TBI).
The impracticality of manually quantifying these lesions limits their
clinical utility and motivates the utilization of machine learning
techniques for automated segmentation workflows.  Herein, we develop a
concatenated random forest framework with image features for
segmenting WMHs in a TBI cohort. The framework is built upon the
Advanced Normalization Tools (ANTs)  and ANTsR toolkits.  MR
(3D FLAIR, T2-, and T1-weighted) images from 24 service members and
veterans scanned in the Chronic Effects of Neurotrauma Consortium’s
(CENC) observational study were acquired. Manual annotations were
employed for both training and evaluation using a leave-one-out
strategy.  Performance measures include sensitivity,
positive predictive value, $F_1$ score, and relative volume difference.
Final average results were: sensitivity = 0.68 $\pm$ 0.38, positive
predictive value = 0.51 $\pm$ 0.40, $F_1$ = 0.52 $\pm$ 0.36,
relative volume difference} = 43 $\pm$ 26\%.  In addition, three
lesion size ranges are selected to illustrate the variation in
performance with lesion size.  Paired with
correlative outcome data, supervised learning methods may allow for
identification of imaging features predictive of diagnosis and prognosis
in individual TBI patients.

>>>>

White matter hyperintensities (WMHs) are foci of abnormal signal
intensity in white matter regions seen with magnetic resonance imaging
(MRI). WMHs are associated with normal aging and have shown prognostic
value in neurological conditions such as traumatic brain injury (TBI).
The impracticality of manually quantifying these lesions limits their
clinical utility and motivates the utilization of machine learning
techniques for automated segmentation workflows.  Herein, we develop a
concatenated random forest framework with image features for
segmenting WMHs in a TBI cohort. The framework is built upon the
Advanced Normalization Tools (ANTs)  and ANTsR toolkits.  MR
(3D FLAIR, T2-, and T1-weighted) images from 24 service members and
veterans scanned in the Chronic Effects of Neurotrauma Consortium’s
(CENC) observational study were acquired. Manual annotations were
employed for both training and evaluation using a leave-one-out
strategy.  Performance measures include sensitivity,
positive predictive value, $F_1$ score, and relative volume difference.
Final average results were: sensitivity = 0.68 $\pm$ 0.38, positive
predictive value = 0.51 $\pm$ 0.40, $F_1$ = 0.52 $\pm$ 0.36,
relative volume difference = 43 $\pm$ 26\%.  In addition, three
lesion size ranges are selected to illustrate the variation in
performance with lesion size.  Paired with
correlative outcome data, supervised learning methods may allow for
identification of imaging features predictive of diagnosis and prognosis
in individual TBI patients.

### 2016-07-07 discussion.Rmd
ENH:  Second round.

In evaluating the segmentation results, there are limited methodologies described in the
literature for automatic segmentation of WMHs despite the numerous clinical studies
[@Debette:2010aa;@Kloppenborg:2014aa] exploring the connections between WMHs and TBI.
However, we can extrapolate from similar application areas where such methodologies have a
much longer history of development such as white matter lesion segmentation in
multiple sclerosis (MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
were found during a literature search and evaluated which concern automated segmentation of white
matter lesions in MS.  The most widely-used data for algorithmic evaluation consists
of two cohorts from two different sites used in
the MS Lesion segmentation challenge associated with the international MICCAI 2008
conference.  For comparison, this challenge training data set
has a mean lesion load of 204 ($\pm$ 752) $mm^3$ per lesion (compared with
our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 $mm^3$ per voxel =
40.56 ($\pm$ 66.48) $mm^3$ per lesion) and the resolution is
almost twice what is used in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).
Note that our relative volume difference numbers are comparable to the
relative volume difference _between raters_ of 68 percent (cf Figure 5 (d))
[@vanginneken:2007aa].

@@

The current communications describes a supervised statistical learning
methodology for identifying WHMs within multimodal MR brain imaging.
This effort utilized information acquired from the manual segmentation
of WMHs from FLAIR images to help build two-stage ensembles of decision
trees for the automated identification of these lesions. Although only a
single expert was used to produce the manual labelings, our intent is to
further refine the proposed paradigm by crowdsourcing with feedback from
other experts who interact with both the data and methodology.  Also, we
recognize that only a single site was used for evaluating the proposed
framework.  However, we are currently processing other site data with
the models developed for this work and the results look promising since
the developed features are site-agnostic.

@@

Two major meta-analyses of WMHs have been published covering the periods
prior to [@Debette:2010aa] and after 2010 [@Kloppenborg:2014aa].
Debette & Markus [@Debette:2010aa] found that the presence of
WMHs was related to subsequent cognitive decline, a higher risk of
developing dementia, stroke, and of mortality. Lesion volume at baseline
was also predictive of cognitive decline. Kloppenborg et al.
[@Kloppenborg:2014aa] of 23 cross-sectional studies reporting MRI and
concurrent neuropsychological results in patients with heterogeneous
diagnoses but without previously diagnosed cognitive impairment, found
that WMHs were associated with cognitive deficit (effect size of -0.10,
95% CI: -0.13 to -0.08) after controlling for age.

>>>>

In evaluating the segmentation results, there are limited methodologies described in the
literature for automatic segmentation of WMHs despite the numerous clinical studies
[@Debette:2010aa;@Kloppenborg:2014aa] exploring the connections between WMHs and TBI.
However, we can extrapolate from similar application areas where such methodologies have a
much longer history of development such as white matter lesion segmentation in
multiple sclerosis (MS).  In a recent literature review [@Garcia-Lorenzo:2013aa], 47 papers
involving  automated segmentation of white
matter lesions in MS.  The most widely-used data for algorithmic evaluation consists
of two cohorts from two different sites used in
the MS Lesion segmentation challenge associated with the international MICCAI 2008
conference.  For comparison, this challenge training data set
has a mean lesion load of 204 ($\pm$ 752) $mm^3$ per lesion (compared with
our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 $mm^3$ per voxel =
40.56 ($\pm$ 66.48) $mm^3$ per lesion) and the resolution is
almost twice what is used in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).
Note that our relative volume difference numbers are comparable to the
relative volume difference _between raters_ of 68 percent (cf Figure 5 (d))
[@vanginneken:2007aa].

@@

The current communications describes a supervised statistical learning
methodology for identifying WMHs within multimodal MR brain imaging.
This effort utilized information acquired from the manual segmentation
of WMHs from FLAIR images to help build two-stage ensembles of decision
trees for the automated identification of these lesions. Although only a
single expert was used to produce the manual labelings, our intent is to
further refine the proposed paradigm by crowdsourcing with feedback from
other experts who interact with both the data and methodology.  Also, we
recognize that only a single site was used for evaluating the proposed
framework.  However, we are currently processing other site data with
the models developed for this work and the results look promising since
the developed features are site-agnostic.

@@

Two major meta-analyses of WMHs have been published covering the periods
prior to [@Debette:2010aa] and after 2010 [@Kloppenborg:2014aa].
Debette & Markus [@Debette:2010aa] found that the presence of
WMHs was related to subsequent cognitive decline, a higher risk of
developing dementia, stroke, and of mortality. Lesion volume at baseline
was also predictive of cognitive decline. Kloppenborg et al.
[@Kloppenborg:2014aa] reviewed 23 cross-sectional studies reporting MRI and
concurrent neuropsychological results in patients with heterogeneous
diagnoses but without previously diagnosed cognitive impairment.  This review found
that WMHs were associated with cognitive deficit (effect size of -0.10,
95% CI: -0.13 to -0.08) after controlling for age.

### 2016-07-07 discussionX.Rmd
ENH:  Second round.

\textcolor{blue}{
In evaluating the segmentation results, there are limited methodologies described in the
literature for automatic segmentation of WMHs despite the numerous clinical studies}
[@Debette:2010aa;@Kloppenborg:2014aa] \textcolor{blue}{exploring the connections between WMHs and TBI.
However, we can extrapolate from similar application areas where such methodologies have a
much longer history of development such as white matter lesion segmentation in
multiple sclerosis (MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
were found during a literature search and evaluated which concern automated segmentation of white
matter lesions in MS.  The most widely-used data for algorithmic evaluation consists
of two cohorts from two different sites used in
the MS Lesion segmentation challenge associated with the international MICCAI 2008
conference.  For comparison, this challenge training data set
has a mean lesion load of 204 ($\pm$ 752) $mm^3$ per lesion (compared with
our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 $mm^3$ per voxel =
40.56 ($\pm$ 66.48) $mm^3$ per lesion) and the resolution is
almost twice what is used in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).
Note that our relative volume difference numbers are comparable to the
relative volume difference between raters of 68 percent (cf Figure 5 (d))}
[@vanginneken:2007aa].

>>>>

\textcolor{blue}{
In evaluating the segmentation results, there are limited methodologies described in the
literature for automatic segmentation of WMHs despite the numerous clinical studies}
[@Debette:2010aa;@Kloppenborg:2014aa] \textcolor{blue}{exploring the connections between WMHs and TBI.
However, we can extrapolate from similar application areas where such methodologies have a
much longer history of development such as white matter lesion segmentation in
multiple sclerosis (MS).  In a recent literature review [@Garcia-Lorenzo:2013aa], 47 papers
involving automated segmentation of white
matter lesions in MS.  The most widely-used data for algorithmic evaluation consists
of two cohorts from two different sites used in
the MS Lesion segmentation challenge associated with the international MICCAI 2008
conference.  For comparison, this challenge training data set
has a mean lesion load of 204 ($\pm$ 752) $mm^3$ per lesion (compared with
our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 $mm^3$ per voxel =
40.56 ($\pm$ 66.48) $mm^3$ per lesion) and the resolution is
almost twice what is used in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).
Note that our relative volume difference numbers are comparable to the
relative volume difference between raters of 68 percent (cf Figure 5 (d))}
[@vanginneken:2007aa].
